---
title: "Healthier U"
author: "Isaac Zhao"
date: "November 26, 2017"
output: html_document
---
```{r}
install.packages("XML")
install.packages("RCurl")
```

```{r}
height <- 65
weight <- 130
```

Create function called Health Analysis. In this function, we will take health variables such as Height and Weight to establish a diagnosis of their BMI. We will write an equation for BMI. No web scraping will be needed.
```{r}
health.analysis <- function(height,weight) {    # Input height (in) and weight (lb) into function

if (height < 2.5) {     # Checks to see if height is impossibly low
    print("Invalid input. Height must be in inches")
} else if (height > 2.5 && weight < 60) {       # Checks to see if weight is impossibly low
    print("Invalid input. Weight must be in pounds")
}
    
BMI <- weight / height^2 * 703      # Converts height (in) and weight (lb) into BMI

# Classifies Diagnosis based on BMI 
if (BMI < 18.5) {
    Diagnosis <- "Underweight"
} else if (BMI >= 18.5 && BMI < 25) {
    Diagnosis <- "Normal Weight"
} else if (BMI >= 25 && BMI < 30) {
    Diagnosis <- "Overweight"
} else if (BMI >= 30) {
    Diagnosis <- "Obese"
}

output<-list(BMI=BMI,Diagnosis=Diagnosis)       # Creates a list so function can return two variables
return(output)      # Function outputs variable BMI and Diagnosis
}

health.analysis(height,weight)
```


Create Function called PopGraph This function will graph your statistics in comparison to the rest of the U.S. population and print out the diagnosis of underweight, normal, or obese. We will conduct web scraping and use ggplot for graphing.
```{r}
library(tidyverse)

popgraph <- function(height,weight) {
    
bmi <- health.analysis(height,weight)$BMI
diagnosis <- health.analysis(height,weight)$Diagnosis
percent <- round(pnorm(bmi,25.6,4) * 100, 2)

ggplot(data.frame(x=c(10,41.2)), aes(x)) +    # Creates plot from x = 10.6 to 40.6
# Plots a normal curve with mean = 25.6, sd = 4. Colors area blue.
    stat_function(fun=dnorm, args=list(25.6,4), color = "dodgerblue", size = 2, geom="area", fill="cadetblue1", alpha = 0.4) +     
    scale_x_continuous(name="BMI") +        # Labels x axis "BMI"
    ggtitle("Percent of United States Population Less Than Given BMI") +   # Adds a graph title
    theme_bw() +        # Makes the background white theme
# Shades only below given BMI dark blue
    stat_function(fun=dnorm, args=list(25.6,4), xlim=c(10,bmi), geom="area", fill="cadetblue3", alpha = 0.7) +   
    
    geom_vline(xintercept=c(18.5,25,30)) +      # Adds black vertical lines in desired locations
    geom_label(aes(35.5,0.08,label=paste0("Your BMI:  ", round(bmi,digits=1), "\n", "Diagnosis:  ", diagnosis)), size=4.3)  +    # prints rounded BMI to 1 decimal and Diagnosis on top right 
    geom_label(aes(35.5,0.0675, label=paste0("Note: BMI may be a misinformative measure", "\n", " of health as it doesn't take into account", "\n", " for muscle mass or body shape.")), color="red", size=2.3) +
    geom_text(aes(25.6,0.02,label=paste0(percent,"%")), size=10)  +     # Adds % label in middle of graph
    geom_text(aes(25.6,0.0001, label=paste0("|", "\n", "US Average")), color="green") +  # Adds average tick mark
    geom_text(aes(15,0.1, label="Underweight")) +
    geom_text(aes(22,0.1, label="Normal Weight")) +
    geom_text(aes(27.5,0.1, label="Overweight")) +
    geom_text(aes(32,0.1, label="Obese")) +
    theme(axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank())    # Removes y axis information
}

popgraph(height,weight)
```


Create Function called Personal Trainer. For this function, we will develop a personalized exercise regimen based off of your current weight, your target weight, and the period from now until the target date. We will incorporate the intensity of the exercise and the number of hours desired to exercise weekly. This will in turn ouput a list of activities. This function only creates a plan for WEIGHT LOSS. We are planning on webscraping data from: http://www.nutristrategy.com/caloriesburned.htm

```{r}
#Webscraping
library(XML)
library(RCurl)
#Save URL into a variable
url.1<-"http://www.nutristrategy.com/caloriesburned.htm"
#Use getURL to download the URL
urldata.1<-getURL(url.1)
#read the table from the website 
data_table.1<-readHTMLTable(urldata.1, stringsAsFactors=FALSE)
#print out data table
data_table.1

data_table.1<-cbind(data_table.1, c(1:249))



```

```{r}
weight <- 130
target.weight <- 120
target.date <- 12       # Period from now until target date in weeks
intensity <- 3     # Total number of hours per week individual can devote to weight loss program

personal.trainer <- function(weight,target.weight,target.date,intensity) {

if (target.weight > weight) {       # Checks to see if target weight is less than current weight
    print("This is a weight loss program")
}

weight.per.week <- (target.weight - weight) / 12    # How many lbs should be lost per week on average

if (weight.per.week >=5){       # Checks to see if weight loss plan is too extreme
    print("Workout plan may be unrealistic and unhealthy")
}

}

personal.trainer(weight,target.weight,target.date,intensity)
```


